### 2022-03-18 outputs/paper/ray.Rmd
some graphs

---
title: "ray"
author: "Ray Wen"
date: "3/18/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(tidyverse)
```

Load Data
```{r}
raw_data <- read.csv("../data/raw_gss.csv")
```

plot
```{r}
# Mental Health Male vs Female

```